Resurgence of the Airy function \\ and other exponential integrals

### 2023-01-20 exp-resurgence.tex
updates

\item Nevanlinna's improvement of Watson's theorem (Sokal, ``An improvement of Watson's theorem on Borel summability'') tells us that for a function $\Phi$ with a well-defined inverse Laplace transform $\phi$, the following conditions are equivalent \textcolor{magenta}{[double check]}:
\item $\Phi$ is Borel regular.
\item $\Phi$ is approximated well, in a certain sense (a weak version of being ``Gevrey-asymptotic''), by its asymptotic series.
\item $\phi$ grows at most exponentially along the Laplace transform ray.
\item When its domain is restricted to the space of Gevrey power series, Borel summation is invertible. \textcolor{magenta}{[Need to find a good reference for this, because a small change in the statement would mess up our conclusion. Does \texttt{arXiv:2112.08792}, \S A.2 have the statement we want?]} Thus, Borel regularization acts as a projection operator on the space of Borel-regularizable functions.\textcolor{orange}{[Veronica: Do you mean that if $\Phi\sim\tilde{\Phi}$ and $\tilde{\Phi}$ is Gevrey and Borel summable, then $\Phi=\mathcal{L}\circ\mathcal{B}\tilde{\Phi}$ (i.e. it is Borel regular)? P.Ramis proved that if $\Phi$ is Gevrey asymptotic to $\tilde{\Phi}$, then $\tilde{\Phi}$ is of Gevrey type. In addition, every Gevrey type series $\tilde{\Phi}$ is the Gevrey asymptotics of a holomorphic function. When it is defined, the Borel sum of $\tilde{\Phi}$ is the natural holomorphic function asymptotic to $\tilde{\Phi}$.]}
\item Borel regularity can help explain why Borel summation is so effective at solving certain kinds of problems, like the ones in Section~\ref{intro:summation}.
\item The formal solutions of a problem are typically supposed to generalize the actual solutions, in the sense that they include the asymptotic series of the actual solutions.
\item In this case, every Borel regular solution can be found by taking the Borel sum of a formal solution.
\item \ldots
\subsection{Goals}

@@

\color{orange}
\begin{itemize}
\item what does it mean being Borel regular?
\item when does it happen?
\begin{itemize}
\item State new Borel regularity results
\begin{itemize}
\item Linear, homogeneous ODE with regular singularity at 0 and irregular singularity at infinity [big idea in {\tt airy-resurgence}]
\begin{itemize}
\item Contextualize with previous work of Braaksma (``Multisummability and Stokes multipliers of linear meromorphic differential equations'')
\item Also contextualize with Balser, Braaksma, Sibuya, and Ramis (``Multisummability of formal power series solutions of linear ordinary differential equations'')
\item Also contextualize with Loday-Richaud
\end{itemize}
\item \emph{Borel regularity} for \textbf{thimbles integrals} can be stated a the commutativity of the following diagram:
\begin{equation}
\begin{tikzcd}
I_{\alpha}(z)\defeq\int_{\mathcal{C}_\alpha}e^{-zf}\nu \arrow[r,"\sim"]\arrow[dd, swap, "\mathcal{L}^\theta "] & \tilde{I}_{\alpha}(z)\arrow[dd,"\mathcal{B}"]\\
& \\
\hat{\iota}_\alpha(\zeta)\arrow[r,equal,swap, "\text{sum}"] & \tilde{\iota}_{\alpha}(\zeta) 
\end{tikzcd}
\end{equation}
\item A priori, the Laplace transform of $\hat{\iota}_\alpha(\zeta)$ and $I_{\alpha}(z)$ have the same asymptotic behaviour in a given sector (indeed taking the asymptotic of $I_\alpha(z)$ we \textit{loose} information); however Borel regularity guarantees that $I_{\alpha}(z)=\mathcal{L}^{\theta}\hat{\iota}_{\alpha}$ in a given sector.
\item fractional derivative formula
\item Conjecturally, we expect $\hat{\varphi}_\alpha(\zeta)$ to have simple singularities. 
\item in the examples, $\hat{\varphi}_\alpha(\zeta)$ turn out be an hypergeometric function of type ${}_pF_{p-1}$ where $p$ is the number of critical values. 
\item We expect that hypergeometric functions play a special role in resurgence theory as they may always appear when there are only finitely many singularities.
\item \textcolor{magenta}{maybe we can say more about algebraic hypergeomtric functions}
\end{itemize}
\item Recall Watson condition (old): Let $R_N$ be the difference between a function and the partial sum
\[ \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \ldots + \frac{\varphi_{N-2}}{z^{N-1}} \]
of its asymptotic series. Watson showed a century ago that the function is Borel regular whenever there's a constant $c \in (0, \infty)$ with
\[ |R_N| \le \frac{c^{N+1} N!}{|z|^N} \]
over all orders $N$ and all $z$ in a wide enough wedge around infinity.
\end{itemize}
\end{itemize}
\color{DarkBlue}
\subsection{Why does Borel resummation work?}
Borel resummation is a way of turning a formal power series
\[ \series{\varphi} = z^\sigma \left( \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \frac{\varphi_3}{z^4} + \ldots \right), \]
with $\sigma \in [0, 1)$, into a function which is asymptotic to $\series{\varphi}$ as $z \to \infty$. Different functions can be asymptotic to the same power series, and Borel resummation picks one of them, performing an implicit regularization~\textbf{[arXiv:1705.03071, or maybe arXiv:1412.6614]}. When a function matches the Borel sum of its asymptotic series, we'll say it's {\em Borel regular}. Several familiar kinds of regularity imply Borel regularity, and shed light on why it occurs.
%%Knowing that a function is Borel regular gives us extra information about it---enough to reconstruct it from its asymptotic series. What's the nature of this extra information?
%%Since different functions can be asymptotic to the same power series, Borel resummation must involve an {\em implicit regularization}, restricting its range to a class of functions which are uniquely determined by their formal power series.
\begin{itemize}
\item \textbf{Having a good asymptotic approximation}

Let $R_N$ be the difference between a function and the partial sum
\[ \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \ldots + \frac{\varphi_{N-2}}{z^{N-1}} \]
of its asymptotic series. Watson showed a century ago that the function is Borel regular whenever there's a constant $c \in (0, \infty)$ with
\[ |R_N| \le \frac{c^{N+1} N!}{|z|^N} \]
over all orders $N$ and all $z$ in a wide enough wedge around infinity (Sokal, ``An improvement of Watson's theorem on Borel summability''; Hardy, {\em Divergent Series}, Theorem~136; Watson, ``A theory of asymptotic series,'' \S 8?).
\item \textbf{Satisfying a singular differential equation}

\begin{itemize}
\item Think about conditions where this works.
\item Maybe the correct place is the setting of Ecalle's formal integral. See \S 5.2.2.1 of Delabaere's {\em Divergent Series, Summability and Resurgence III}.
\item Say there's a unique solution (up to scaling) that shrinks as you go right; everything else blows up exponentially. Then this is the only solution that can be expressed as a Laplace transform.
\item If the Borel-transformed equation has a subexponential solution $\hat{f}$ which is ``shifted holomorphic'' (we called this having a ``fractional power singularity'' in {\tt airy-resurgence}), then $\laplace \hat{f}$ satisfies the original equation, because there are no boundary terms.
\item Draw diagram showing formal vs. holomorphic solutions in time vs. frequency domains.
\end{itemize}
\item \textbf{Being a thimble integral}

>>>>

\item Borel regular functions have been characterized in the literature before; Watson showed a century ago \textcolor{orange}{[Watson]} that a function $\Phi$ is Borel regular whenever there's a constant $c \in (0, \infty)$ with
\[ |\Phi(z)-\sum_{k=0}^Nc_kz^{-k}| \le c^{N+1} N!|z|^{-N} \]
over all orders $N$ and all $z$ in a wide enough wedge around infinity.

Nevanlinna's improvement of Watson's theorem (Sokal, ``An improvement of Watson's theorem on Borel summability'') tells us that for a function $\Phi$ with a well-defined inverse Laplace transform $\phi$, the following conditions are equivalent \textcolor{magenta}{[double check]}:
\begin{itemize}
\item $\Phi$ is Borel regular.
\item $\Phi$ is approximated well, in a certain sense (a weak version of being ``Gevrey-asymptotic''), by its asymptotic series.
\item $\phi$ grows at most exponentially along the Laplace transform ray.
\end{itemize}
\item When its domain is restricted to the space of Gevrey power series, Borel summation is invertible. \textcolor{magenta}{[Need to find a good reference for this, because a small change in the statement would mess up our conclusion. Does \texttt{arXiv:2112.08792}, \S A.2 have the statement we want?]} Thus, Borel regularization acts as a projection operator on the space of Borel-regularizable functions.\textcolor{orange}{[Veronica: Do you mean that if $\Phi\sim\tilde{\Phi}$ and $\tilde{\Phi}$ is Gevrey and Borel summable, then $\Phi=\mathcal{L}\circ\mathcal{B}\tilde{\Phi}$ (i.e. it is Borel regular)? P.Ramis proved that if $\Phi$ is Gevrey asymptotic to $\tilde{\Phi}$, then $\tilde{\Phi}$ is of Gevrey type. In addition, every Gevrey type series $\tilde{\Phi}$ is the Gevrey asymptotics of a holomorphic function. When it is defined, the Borel sum of $\tilde{\Phi}$ is the natural holomorphic function asymptotic to $\tilde{\Phi}$.]}
\item Borel regularity can help explain why Borel summation is so effective at solving certain kinds of problems, like the ones in Section~\ref{intro:summation}.
\item The formal solutions of a problem are typically supposed to generalize the actual solutions, in the sense that they include the asymptotic series of the actual solutions.
\item In this case, every Borel regular solution can be found by taking the Borel sum of a formal solution.
\item \ldots
\end{itemize}
\subsection{Goals}

@@

\color{orange}
\begin{itemize}
\item what does it mean being Borel regular?
\item when does it happen?
\begin{itemize}
\item State new Borel regularity results
\begin{itemize}
\item Linear, homogeneous ODE with regular singularity at 0 and irregular singularity at infinity [big idea in {\tt airy-resurgence}]
\begin{itemize}
\item Contextualize with previous work of Braaksma (``Multisummability and Stokes multipliers of linear meromorphic differential equations'')
\item Also contextualize with Balser, Braaksma, Sibuya, and Ramis (``Multisummability of formal power series solutions of linear ordinary differential equations'')
\item Also contextualize with Loday-Richaud
\end{itemize}
\item \emph{Borel regularity} for \textbf{thimbles integrals} can be stated a the commutativity of the following diagram:
\begin{equation}
\begin{tikzcd}
I_{\alpha}(z)\defeq\int_{\mathcal{C}_\alpha}e^{-zf}\nu \arrow[r,"\sim"]\arrow[dd, swap, "\mathcal{L}^\theta "] & \tilde{I}_{\alpha}(z)\arrow[dd,"\mathcal{B}"]\\
& \\
\hat{\iota}_\alpha(\zeta)\arrow[r,equal,swap, "\text{sum}"] & \tilde{\iota}_{\alpha}(\zeta) 
\end{tikzcd}
\end{equation}
\item A priori, the Laplace transform of $\hat{\iota}_\alpha(\zeta)$ and $I_{\alpha}(z)$ have the same asymptotic behaviour in a given sector (indeed taking the asymptotic of $I_\alpha(z)$ we \textit{loose} information); however Borel regularity guarantees that $I_{\alpha}(z)=\mathcal{L}^{\theta}\hat{\iota}_{\alpha}$ in a given sector.
\item fractional derivative formula
\item Conjecturally, we expect $\hat{\varphi}_\alpha(\zeta)$ to have simple singularities. 
\item in the examples, $\hat{\varphi}_\alpha(\zeta)$ turn out be an hypergeometric function of type ${}_pF_{p-1}$ where $p$ is the number of critical values. 
\item We expect that hypergeometric functions play a special role in resurgence theory as they may always appear when there are only finitely many singularities.
\item \textcolor{magenta}{maybe we can say more about algebraic hypergeomtric functions}
\end{itemize}
\item Recall Watson condition (old): Let $R_N$ be the difference between a function and the partial sum
\[ \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \ldots + \frac{\varphi_{N-2}}{z^{N-1}} \]
of its asymptotic series. Watson showed a century ago that the function is Borel regular whenever there's a constant $c \in (0, \infty)$ with
\[ |R_N| \le \frac{c^{N+1} N!}{|z|^N} \]
over all orders $N$ and all $z$ in a wide enough wedge around infinity.
\end{itemize}
\end{itemize}
\color{DarkBlue}
\subsection{Why does Borel resummation work?}

\color{gray}

Borel resummation is a way of turning a formal power series
\[ \series{\varphi} = z^\sigma \left( \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \frac{\varphi_3}{z^4} + \ldots \right), \]
with $\sigma \in [0, 1)$, into a function which is asymptotic to $\series{\varphi}$ as $z \to \infty$. Different functions can be asymptotic to the same power series, and Borel resummation picks one of them, performing an implicit regularization~\textbf{[arXiv:1705.03071, or maybe arXiv:1412.6614]}. When a function matches the Borel sum of its asymptotic series, we'll say it's {\em Borel regular}. Several familiar kinds of regularity imply Borel regularity, and shed light on why it occurs.
%%Knowing that a function is Borel regular gives us extra information about it---enough to reconstruct it from its asymptotic series. What's the nature of this extra information?
%%Since different functions can be asymptotic to the same power series, Borel resummation must involve an {\em implicit regularization}, restricting its range to a class of functions which are uniquely determined by their formal power series.
\begin{itemize}
\item \textbf{Having a good asymptotic approximation}

Let $R_N$ be the difference between a function and the partial sum
\[ \frac{\varphi_0}{z} + \frac{\varphi_1}{z^2} + \frac{\varphi_2}{z^3} + \ldots + \frac{\varphi_{N-2}}{z^{N-1}} \]
of its asymptotic series. Watson showed a century ago that the function is Borel regular whenever there's a constant $c \in (0, \infty)$ with
\[ |R_N| \le \frac{c^{N+1} N!}{|z|^N} \]
over all orders $N$ and all $z$ in a wide enough wedge around infinity (Sokal, ``An improvement of Watson's theorem on Borel summability''; Hardy, {\em Divergent Series}, Theorem~136; Watson, ``A theory of asymptotic series,'' \S 8?).
\color{DarkBlue}

\item \textbf{Satisfying a singular differential equation}

\begin{itemize}
\item Think about conditions where this works.
\item Maybe the correct place is the setting of Ecalle's formal integral. See \S 5.2.2.1 of Delabaere's {\em Divergent Series, Summability and Resurgence III}.
\item Say there's a unique solution (up to scaling) that shrinks as you go right; everything else blows up exponentially. Then this is the only solution that can be expressed as a Laplace transform.
\item If the Borel-transformed equation has a subexponential solution $\hat{f}$ which is ``shifted holomorphic'' (we called this having a ``fractional power singularity'' in {\tt airy-resurgence}), then $\laplace \hat{f}$ satisfies the original equation, because there are no boundary terms.
\item Draw diagram showing formal vs. holomorphic solutions in time vs. frequency domains.

\begin{center}
\begin{tikzcd}
& & ODE\arrow[dd,"\mathcal{B}"]\arrow[dd,"\mathcal{L}^{-1}",swap] & & \\
\Phi\arrow[rru,dashed] &\textcolor{red}{\hat{\Phi}}\arrow[ur,red] & & & \tilde{\Phi}\arrow[llu,"formal",swap]\arrow[dd,"\mathcal{B}"]\\
& & IE & & \\
\phi\arrow[rru]\arrow[uu,"\mathcal{L}"]& \textcolor{red}{\hat{\phi}}\arrow[ur]\arrow[rrr,"sum",mapsfrom,red]\arrow[uu,"\mathcal{L}"]& & & \tilde{\phi}\arrow[llu]
\end{tikzcd}
\end{center}

where the arrow in red are a consequence of multisummability. In addition, we can distinguish on the right hand side of the diagram the formal solutions and on the left hand side the holomorphic ones. On the upper part of the diagram the functions in the $z$-plane while on the lower part the functions in the Borel plane $\zeta$-plane. 
\end{itemize}

\item \textbf{Being a thimble integral}
